---
title: "A meta-research on the Cochrane Reviews' authors"
author: "Ahmad Sofi-Mahmudi"
date: "`r Sys.Date()`"
output: html_document
---

# Loading required packages

```{r}
pacman::p_load(dplyr,
               here,
               knitr,
               ggplot2,
               maps,
               stringr,
               rnaturalearth)
```


# Loading the dataset
```{r}
db <- read.csv(here("data", "cochraneauthors_db.csv"), 
               header = TRUE,
               na.strings = c("", "NA"))
```

# Overall perspective

Total number of papers:
```{r}
nrow(reviews)
```

Number of papers that their authors affiliations were retrieved from Cochrane Library successfully:
```{r}
nrow(db)
```

Percentage of affiliations successfully retrieved:
```{r}
round(nrow(db)/8824*100, 1)
```


# Country and region diversity

## First authors
### Countries

The number of countries detected successfully:
```{r}
sum(table(db$first_author_country))
```

Top ten countries with the highest number of first authors:
```{r}
top_first <- table(db$first_author_country) %>% 
  as.data.frame() %>%
  arrange(desc(Freq))

colnames(top_first) <- c("country", "frequency")

kable(top_first[1:10,])
```

Map
```{r}
# Loading the world map
world_map <- map_data("world")
world_map <- subset(world_map, region != "Antarctica")

# Some modifications are needed
top_first$country <- as.character(top_first$country)
top_first$country[top_first$country == "United States"] <- "USA"
top_first$country[top_first$country == "United Kingdom"] <- "UK"
top_first$country <- as.factor(top_first$country)

# Drawing the map
ggplot(top_first) +
  geom_map(
    dat = world_map, map = world_map, aes(map_id = region),
    fill = "white", color = "#7f7f7f", size = 0.25
  ) +
  geom_map(map = world_map, aes(map_id = country, fill = frequency), size = 0.25) +
  scale_fill_gradient(low = "#fff7bc", high = "#cc4c02", name = "Total First Authors") +
  expand_limits(x = world_map$long, y = world_map$lat) + theme(legend.position="bottom",
        axis.line=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        panel.background=element_blank(),
        panel.border=element_blank(),
        panel.grid=element_blank())
```


### United Nations regions
The number of countries that their region detected successfully:
```{r}
sum(table(db$first_author_region_un))
```

Top regions with the highest number of first authors:
```{r}
top_un_first <- table(db$first_author_region_un) %>% 
  as.data.frame() %>%
  arrange(desc(Freq))

colnames(top_un_first) <- c("un region", "frequency")

kable(top_un_first)
```

### World Bank regions
The number of countries that their region detected successfully:
```{r}
sum(table(db$first_author_region_wb))
```

Top regions with the highest number of first authors:
```{r}
top_wb_first <- table(db$first_author_region_wb) %>% 
  as.data.frame() %>%
  arrange(desc(Freq))

colnames(top_wb_first) <- c("wb region", "frequency")

kable(top_wb_first)
```
### Income-level regions
The number of countries that their region detected successfully:
```{r}
sum(table(db$first_author_region_income))
```

Top regions with the highest number of first authors:
```{r}
top_income_first <- table(db$first_author_region_income) %>% 
  as.data.frame() %>%
  arrange(desc(Freq))

colnames(top_income_first) <- c("income region", "frequency")

kable(top_income_first)
```

## Last authors
### Countries

### United Nations regions

### World Bank regions

### Income-level regions


## Corresponding authors
### Countries

### United Nations regions

### World Bank regions

### Income-level regions


# English-speaking diversity

## Trend
```{r}

yearly_trend <- db %>%
  select(year, first_author_english) %>%
  group_by(year) %>%
  summarise(counts = n(),
            english_speaking = sum(first_author_english == "Yes")) %>%
  mutate(p = english_speaking / counts)


p1 <-
        yearly_trend %>% 
        ggplot() +
        aes(x = year, 
            y = p) +
        geom_line(size = 0.75) +
        scale_y_continuous(limits = c(0, 1), 
                           labels = scales::percent) +
        scale_color_discrete(name = NULL) +
        labs(   title="Proportion of papers with first authors from developed English-speaking countries",
                y = "Proportion of articles (%)", 
                x = "Year") +
        theme(panel.grid.minor = element_blank(),
              legend.position = c(0.13, 0.87),
              axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)
              )
p1
```


# Gender diversity

### Trend
```{r}

yearly_trend_gender <- db %>%
  select(year, country_1stauthor_gender) %>%
  group_by(year) %>%
  summarise(MtoF = sum(country_1stauthor_gender == "M")/sum(country_1stauthor_gender == "F"))


pg1 <-
        yearly_trend_gender %>% 
        ggplot() +
        aes(x = year, 
            y = MtoF) +
        geom_line(size = 0.75) +
        scale_color_discrete(name = NULL) +
        labs(   title="Male to female ratio for each year",
                y = "Ratio", 
                x = "Year") +
        theme(panel.grid.minor = element_blank(),
              legend.position = c(0.13, 0.87),
              axis.text.x = element_text(angle = 45, vjust = 0.9, hjust=1)
              )
pg1
```

